\documentclass{article}

\author{Kody Manastyrski}
\date{January 2022}
\title{CP 8319 Assignment 1}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}

\begin{document}

\maketitle

\section{Question 1}
\subsection{A}
The following matrix shows the values of $r_s$, and is treated as a lookup table.
\[
\begin{bmatrix}
	0 & 0 & 0 & 0 \\
	-5 & -1 & 0 & 1 \\
	0 & 0 & 1 & 5 \\
	0 & 0 & 0 & 0
\end{bmatrix}
\]
Note: Here the $r_g$ and $r_s$ values are included.

Below is the values for each square (i.e. shortened form for the answer), 
followed by the work to obtain them.
\[
\begin{bmatrix}
	6 & 6 & 6 & 6 \\
	-5 & 5 & 6 & 6 \\
	6 & 6 & 6 & 6 \\
	6 & 6 & 6 & 6
\end{bmatrix}
\]

\begin{table}[H]
\begin{tabular}{|l|c|r|}
	\hline
	Cell & Values on Path (Self included) & Total\\
	\hline
	1 & 0 + 0 + 0 + 0 + 1 + 5 & 6\\
	2 & 0 + 0 + 0 + 1 + 5 & 6 \\
	3 & 0 + 0 + 1 + 5 & 6 \\
	4 & 0 + 1 + 5 & 6 \\
	5 & -5 & -5 \\
	6 & -1 + 0 + 1 + 5 & 5 \\
	7 & 0 + 1 + 5 & 6 \\
	8 & 1 + 5 & 6 \\
	9 & 0 + 0 + 1 + 5 & 6 \\
	10 & 0 + 1 + 5 & 6 \\
	11 & 1 + 5 & 6 \\
	12 & 0 + 0 + 0 + 1 + 5 & 6 \\
	13 & 0 + 0 + 0 + 0 + 1 + 5 & 6 \\
	14 & 0 + 0 + 0 + 0 + 0 + 1 + 5 & 6 \\
	15 & 0 + 0 + 0 + 0 + 0 + 0 + 1 + 5 & 6 \\
	16 & 0 + 0 + 0 + 0 + 0 + 0 + 0 + 1 + 5 & 6 \\
	\hline
\end{tabular}
\end{table}


\subsection{B}
\[
\begin{bmatrix}
	18 & 16 & 14 & 12 \\
	-3 & 13 & 12 & 10 \\
	14 & 12 & 10 & 7 \\
	16 & 18 & 20 & 22
\end{bmatrix}
\]

\subsection{C}
\[
	V^{\pi}_{new[i,j]} = (c +\gamma V^{\pi}_{old[i,j]}) + \displaystyle\sum^{|\mathfrak{P}|}_{p \in \mathfrak{P}, x = 0} c + \gamma^x V^{\pi}_{old[p]}
\]
Where $[i,j]$ is the indices of the $V^{\pi}_{old}$ table, and $\mathfrak{P}$ is the ordered sequence of indices for the optimal path from the next step in the path starting at $[i,j]$.
For example, with the above new, starting at $[i=0,j=0]$ we have $\mathfrak{P}=
\{[1,0], [2,0], [3,0],[3,1]\}$ assuming indices start at 0, are in (row, column)
order, and grow moving to the right for rows, and down for columns.
Note that the x goes from 0 to the count of elements in the pathset, while p is
just the indices at each xth step.

\subsection{D}
Supposing that the gridworld now has $r_s = 2 \forall r$ then the world changes
to reward the longest path, since all cells have the same reward value.
On the other hand, if the world merely adds 2 to the existing old rewards, then 
the gridworld disincentives the terminal states, and incentives moving away
from the terminals (at least initially).
The values of the shaded squares are then $\{v \in V^{\pi}_{new}\}$, when adding 2 to
any given square, and $2*|\mathfrak{P}|$ when setting all squares to 2.

\subsection{E}
Yes, the optimal policy in fact does change as soon as gamma is not 1, and not 0.
This puts a timer and an upper limit on the point value, so moving fast and 
terminating quickly is incentivised, rather than meandering and accumulating 
further points by lengthening the path.

\subsection{F}
\[ r_s = 
	\begin{bmatrix}
		1 & 2 & 3 & 4 \\
		-5 & 10 & -5 & 5 \\
		10 & -5 & 0 & 5\\
		4 & 3 & 2 & 1
	\end{bmatrix}
\]
This assumes that "there are" refers to the logical "there exists", and thus some
squares terminate at cell 5, and some at cell 12.

\end{document}
